\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-06-23 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates on progress

Equation~\ref{eqn:saturation_eqn} is discretised in time and space using the $\theta$-method. Summing over all phases yields the global mass balance equation, 
\begin{eqnarray}
 && \sum_{\alpha=1}^{N_{p}} \left\{\int\limits_{\Omega_{CVi}} \; M_{i} \; \frac{\phi\left(S_{\alpha i}^{n+1}-S_{\alpha i}^{n}\right)}{\Delta t} dV\right.  + \nonumber\\
 &&  \oint_{\Gamma_{CVi}} \left[\; \theta^{n+1/2}\; {\mathbf n} \cdot \; {\mathbf u}_{\alpha}^{n+1}S_{\alpha}^{n+1} \; + \; \left(1-\theta^{n+1/2}\right) \; {\mathbf u}_{\alpha}^{n}S_{\alpha}^{n}\right] d\Gamma \;- \nonumber\\
 &&  \left.\int_{\Omega_{CVi}} M_{i} \; \mathcal{S}_{cty,\alpha}^{n+\theta} \; dV\;\right\} =0,
\label{global_mass_balance}
\end{eqnarray}
where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, $M_{i}$ are CV basis functions, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $n$ is the current time level. $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. In addition to the numerical schemes above the discretised global mass and force balance equations are solved using a multigrid-like approach, see \citet{pavlidis2016}. This numerical formulation is fully described in \citet{gomes_2017}\citep[see also][]{salinas2015,adam_2016}.

@@

\medskip
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits may cause instabilities in the system. The flow of a fluid over or next to another heavier fluid, can produce flow instabilities. Under this assumption when a fluid is injected in a porous domain saturated with another fluid, viscous fingering may occur, as viscous driven hydrodynamic instability leads to the collapse of the uniform interface between fluids of different viscosity.

>>>>

\medskip
Equation~\ref{eqn:saturation_eqn} is discretised in time and space using the $\theta$-method. Summing over all phases yields the global mass balance equation, 
\begin{eqnarray}
 && \sum_{\alpha=1}^{N_{p}} \left\{\int\limits_{\Omega_{CVi}} \; M_{i} \; \frac{\phi\left(S_{\alpha i}^{n+1}-S_{\alpha i}^{n}\right)}{\Delta t} dV\right.  + \nonumber\\
 &&  \oint_{\Gamma_{CVi}} \left[\; \theta^{n+1/2}\; {\mathbf n} \cdot \; {\mathbf u}_{\alpha}^{n+1}S_{\alpha}^{n+1} \; + \; \left(1-\theta^{n+1/2}\right) \; {\mathbf u}_{\alpha}^{n}S_{\alpha}^{n}\right] d\Gamma \;- \nonumber\\
 &&  \left.\int_{\Omega_{CVi}} M_{i} \; \mathcal{S}_{cty,\alpha}^{n+\theta} \; dV\;\right\} =0,
\label{global_mass_balance}
\end{eqnarray}
where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, $M_{i}$ are CV basis functions, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $n$ is the current time level. $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. In addition to the numerical schemes above the discretised global mass and force balance equations are solved using a multigrid-like approach, see \citet{pavlidis2016}. This numerical formulation is fully described in \citet{gomes_2017}\citep[see also][]{salinas2015,adam_2016}.

@@

\medskip
As in many areas of fluid mechanics, the behaviour of the flow for conditions that exceed critical limits (unstable interface because of viscocity differences, density differences,the existance of gravitational forces or not, and/or opposite stress tensors) may cause instabilities in the system. The flow of a fluid over or next to another heavier fluid, can produce flow instabilities. Under this assumption when a fluid is injected in a porous domain saturated with another fluid, viscous fingering may occur, as viscous driven hydrodynamic instability leads to the collapse of the uniform interface between fluids of different viscosity.
